---
title: "95% Confidence Intervals"
subtitle: "Introduction to Statistical Methods (PALS0046)\ngithub.com/sdparsons/"
author: "Sam Parsons"
date: "25/06/2023"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 95% Confidence Intervals

95 % confidence is a confidence that in the long-run 95 % of the CIs will include the population mean

95\%CI = point estimate +/- critical value*standard error


```{r CI fig, echo = FALSE, message = FALSE, warning=FALSE, fig.show='asis', results='hold'}
trees <- datasets::trees

# simulate population level data
pop_level <- rnorm(n = 100000, 
                   mean = mean(trees$Height), 
                   sd(trees$Height))

# check summary statistics
mean(pop_level)
sd(pop_level)

# take 20 samples
n_samples <- 20
sample_size <- 20

sample_heights <- NULL

for(i in 1:n_samples) {
  sample_temp <- sample(pop_level, sample_size, replace = TRUE)
  sample_heights <- c(sample_heights, sample_temp)
}

sample_data <- data.frame(sample = rep(1:n_samples, each = sample_size),
                          heights = sample_heights)

library(ggplot2)
library(dplyr)
sample_data %>%
  group_by(sample) %>%
  summarise(mean = mean(heights),
            se = sd(heights)/sqrt(n())) %>%
  mutate(lowerCI = mean + (qnorm(.025) * se),
         upperCI = mean + (qnorm(.975) * se)) %>%
  ggplot(aes(x = sample,
             y = mean)) +
  geom_point() +
  geom_abline(intercept = mean(pop_level), slope = 0, colour = "red", linewidth = 1.1) +
  geom_errorbar(aes(ymin = lowerCI,
                ymax = upperCI)) +
  coord_cartesian(ylim = c(68, 83))



```

## What the 95% CI is not

the chance the true value falls within your CI

```{r CI fig2, echo = FALSE, message = FALSE, warning=FALSE, fig.show='asis', results='hold', out.height="40%"}
sample_data %>%
  group_by(sample) %>%
  filter(sample == 12) %>%
  summarise(mean = mean(heights),
            se = sd(heights)/sqrt(n())) %>%
  mutate(lowerCI = mean + (qnorm(.025) * se),
         upperCI = mean + (qnorm(.975) * se)) %>%
  ggplot(aes(x = sample,
             y = mean)) +
  geom_point() +
  geom_abline(intercept = mean(pop_level), slope = 0, colour = "red", linewidth = 1.1) +
  geom_errorbar(aes(ymin = lowerCI,
                ymax = upperCI),
                width = .1) +
  coord_cartesian(xlim = c(11.7, 12.3), ylim = c(68, 83))

```

(we will discuss a bunch of weird statistical things this term, stay tuned)


## How tall are our trees


```{r data}
trees <- datasets::trees

head(trees)

```

## Plotting tree height

```{r hist}
hist(trees$Height)
```

## 95%CI of the mean



```{r ci of mean}
mean(trees$Height)

sd(trees$Height)

se = sd(trees$Height) / sqrt(nrow(trees))

se

```

## 95%CI of the mean

95\%CI = point estimate +/- critical value*standard error

lower and upper critical values

```{r}
qnorm(c(.025,.975))
```

```{r}
mean(trees$Height) + qnorm(.025) * se
mean(trees$Height) + qnorm(.975) * se

```

## Bootstraped Confidence Intervals

Some statistics are more complex, we may not have a closed-form equation

Bootstrapping involves taking lots of random samples from your data

For each re-sample, you calculate your statistic

Then, you compile the whole distribution to extract the confidence intervals

In this example we're interested in the mean tree BMI, for reasons





## Exercises

- calculate 90% CIs for tree height. How do the 90% and 95% CIs differ? what does this mean?
- calculate 95% CIs for tree girth (diameter) and volume

*bootstrapping exercises*

- calculate bootstrapped CIs for tree height
- plot the bootstrapped CIs (check github code for hints)




